IFDS 2021 Spring RA research proposal

### 2021-11-02 IFDS/ifds21Sra_proposal/main.tex
proposal

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

@@

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\usepackage{colonequals}

@@

\title{Future Improvement}

@@

% Abstracts are required.
\section{Unsolved Review Summary}

\subsection{Comparisons}
\begin{enumerate}
	\item Add comparison on tensor-on-tensor regression~\citep{lock2018tensor,llosa2018tensor,gahrooei2020multiple,raskutti2015convex}. {\bf NIPS R3}.
	\item Add comparison on scalar-on-tensor regression~\citep{chen2019non, zhou2013tensor}. {\bf ICML R1}.
	\item Add comparison on HOLRR~\citep{rabusseau2016low}. Though HOLRR is a special case of the proposed model, it is helpful to show the benefit of the incorporation of covariate on distinct modes, when all covariates are vectorized in a long vector in HOLRR. {\bf AISTAT R3}.
	\item Add comparison on HOPLS~\citep{zhao2012higher} and TPG \citep{yu2016learning} from the perspective of the model. {\bf AISTAT R1}.
	\item Add comparison on paper~\citep{smith2015positive}. {\bf NIPS R4}.
\end{enumerate}

\subsection{Experiments}
\begin{enumerate}
	\item Add comparisons on other scalar-on-tensor and tensor-on-tensor regressions in both simulations and real data analysis. {\bf NIPS R2, R3}.
	\item Add numerical results on large-scale or higher-order response. {\bf NIPS R2, AISTAT R3}.
	\item Show how many iterations and time are needed to converge. {\bf NIPS R2}.
\end{enumerate}

\subsection{Explanations}
\begin{enumerate}
	\item Add explanations on hyperparameter selection for our method and other methods (e.g.\ HOLRR, TPG, HOPLS) we want to compare. {\bf NIPS R2,R4}.
	\item Add explanations on rank selection. Add necessary references. Explain why not use the proposed rank selection strategy in simulations. Explain the computational issues of grid search. {\bf NIPS R3, ICML R1, R3}.
	\item Discuss the existence of $\tB_{true}$ and discuss the mismatching loss $\sum \hat \tB_{ijk} \neq {\tB_{true}}_{ijk}$. {\bf NIPS R4}.
	\item Compare the computational complexity with other methods. {\bf NIPS R2}.
	\item Explain how the 136 subjects in HCP are selected from thousands of subjects in the original data. {\bf NIPS R3}.
\end{enumerate}

\section{Model Equivalence}

\subsection{Our method}
Let $\tY = \entry{\tY_{i_1,...,i_k}} \in \bbR^{d_1,...,d_K}$ be the tensor observation, $\mX_k \in \bbR^{d_k \times p_k}, k \in [K]$ be the matrix covariates. Our tensor regression model is of form
\begin{equation}\label{ourmodel}
	g \of{ \bbE[\tY |\mX_1 ,...,\mX_K ] } = \tB \times_1 \mX_1 \times_2 \cdots \times_K \mX_K,
\end{equation}
where $g(\cdot)$ is the inverse link function, $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient. Reformulating the model~\eqref{ourmodel} into $\prod_{i=1}^K d_k$ univariate regressions, we have
\begin{equation}\label{ourmodel1}
	g \of{ \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] }  = \langle \tB, \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}  \rangle, \quad \text{for } i_k \in [d_k], k\in[K],
\end{equation}
where $\mX_{k i_k}$ refers to the $i_k$-th row of $\mX_k$, for all $k \in [K]$, $\langle \cdot, \cdot \rangle$ refers to the inner product of two tensors, and $\circ$ refers to the outer product of two vectors.

\subsection{Scalar-on-tensor regression}
Let $y \in \bbR$ denote the scalar response, $\tX \in \bbR^{p_1 \times \cdots \times p_K}$ denote the order-K tensor predictor, and $\mZ \in \bbR^{p_0}$ denote the vector-valued covariate. Given the i.i.d.\ sample set $\{y^{(i)}, \tX^{(i)}, \mZ^{(i)}\}_{i=1}^n$,  the scalar-on-tensor regression model is
\begin{equation}\label{scalar}
	g \of{\bbE[y^{(i)} | \tX^{(i)}]}  = \alpha + \gamma^T \mZ + \langle \tB, \tX^{(i)}  \rangle, \quad \text{for } i  = 1 ,...,n
\end{equation}

where $g(\cdot)$ is the inverse link function, $\alpha \in \bbR$ is the scalar coefficient, $\gamma \in \bbR^{p_0}$ is the vector-valued coefficient, and $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient.

Ignoring the scalar parameter $\alpha$ and the vector-valued covariate $\mZ$, our model~\eqref{ourmodel1} is equal to the scalar-on-tensor model~\eqref{scalar} by defining
\begin{equation}
	\bbE[y^{(i)} | \tX^{(i)}] \colonequals \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] \quad \text{and} \quad \tX^{(i)} \colonequals \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}, \quad \text{for } i = 1,..., n,
\end{equation}
where $n = \prod_{i=1}^K d_k$.

\subsection{Tensor-on-tensor regression}
Let $\mathbb{Y} \in \bbR^{ d_{K+1} \times \cdots \times d_{N}}$ be the tensor observation, $\mathbb{X} \in \bbR^{ d_1 \times \cdots \times d_K}$ be the tensor predictor. Given the i.i.d.\ sample set $\{\mathbb{Y}^{(i)}, \mathbb{X}^{(i)} \}_{i=1}^n$,  the tensor-on-tensor regression is of form
\begin{equation}\label{tensor}
	g \of{\bbE[\mathbb{Y}^{(i)} | \mathbb{X}^{(i)}]} = \langle \mathbb{X}^{(i)}, \mathbb{B}   \rangle_K, \quad \text{for } i = 1,...,n,
\end{equation}

where $g(\cdot)$ is the inverse link function, $\mathbb{B} \in \bbR^{d_1 \times \cdots \times d_N}$ is the order-K tensor-valued coefficient, $\langle \cdot, \cdot \rangle_K$ refers to the contracted tensor product; i.e., for two tensors $\mathbb{P} \in \bbR^{I_1 \times \cdots \times I_L \times d_1 \times \cdots \times d_K}$ and $\mathbb{Q} \in \bbR^{d_1 \times \cdots \times d_K \times J_1 \times \cdots \times J_M}$, the contracted tensor product $\langle \mathbb{P}, \mathbb{Q} \rangle_K \in \bbR^{I_1\times \cdots \times I_L \times J_1 \times \cdots \times J_M}$ and the $(i_1,...,i_L, j_1,...,j_M)$-the entry is
\begin{equation}
	(\langle \mathbb{P}, \mathbb{Q} \rangle_K)_{i_1,...,i_L, j_1,...,j_M} = \sum_{a_1,...,a_K}^{d_1,...,d_K} \mathbb{P}_{i_1,...,i_L, a_1,...,a_K} \mathbb{Q}_{a_1,...,a_K, j_1,...,j_M}.
\end{equation}

Let $N = K$. Then, the contracted product becomes usual inner product, and $\mathbb{Y}^{(i)}$ becomes a scalar for $i \in [n]$. Consequently, the model~\eqref{tensor} degenerates to the scalar-on-tensor regression~\eqref{scalar}, which is equal to our model.

\section{Iteration time table}

See Table~\ref{tab:time}.
\begin{table}[h]
\centering
%\resizebox{1\textwidth}{!}{
\begin{tabular}{|c|c|c|c|}
\hline
Setting &Gaussian& Bernoulli & Poisson  \\ \hline
$d = 30, r = 6$& 2.6 &  6.0& 5.7\\
$d = 30, r = 3$&  0.5 & 1.1&1.3\\
$d = 25, r = 6$&1.4 & 3.7& 4.0\\
$d = 25, r = 3$&0.3& 0.6& 0.6\\
\hline
\end{tabular}
%}
\caption{ Iteration time of Algorithm 1. Numerical values are the running time (seconds) required for one iteration in Algorithm 1 under different settings.  }\label{tab:time}
\end{table}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

>>>>

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

@@

@@

\setlength{\parskip}{0.8em}

\title{IFDS 2021 Spring RA research proposal}

@@

Multiway arrays are collected in many fields including social networks \citep{anandkumar2014tensor} and computer science \citep{koniusz2016sparse}, and tensors represent the multiway data efficiently. One example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science. A $K$-uniform hypergraph can be naturally represented as an order-$K$ tensor, in which each entry indicates the presence of $K$-way hyperedge among nodes. Identifying the similarity among nodes is of scientific interests.  We refer such clustering as \emph{higher-order clustering}. A huge amounts methods have been developed for matrix clustering while applying matrix methods to tensors results in sub-optimal performances. Unsupervised tensor models and tools for higher-order clustering have arisen recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.

Meanwhile, semi-supervised learning has been widely applied to clustering problems~\citep{zhu2003semi,yang2014unified}. The potential useful prior information such as a small proportion of labeled data and auxiliary features are well-incorporated in the learning tasks, which is believed to improve clustering performance than unsupervised methods. Semi-supervised learning are also used in tensor decomposition~\citep{cao2016semi,hu2021generalized}. However, applying semi-supervised techniques in higher-order clustering is still new to the field. Several questions such as how to incorporate prior information in tensor models and how the prior information affects the clustering performance are waiting to be answered.

Therefore, this project will focus on the application of semi-supervised learning in higher-order clustering problem. The work will enrich the methodology in higher-order clustering literature and deepen the understanding of semi-supervised learning in higher-order regime, which tackles the main theme of IFDS.

\textbf{Bio:}

Jiaxin Hu is a second-year Ph.D.\ student in Statistics at the University of Wisconsin-Madison, advised by Prof. Miaoyan Wang. Her research mainly focuses on the statistical machine learning methods for matrix/tensor data analysis. She actively attended the weekly IFDS seminars and the summer school and workshop at IFDS in August, 2021. During the workshop, she gave a five minute lightening talk with title ``Supervised tensor decomposition with features on multiple modes".

\bibliography{tensor_wang.bib}
\bibliographystyle{apalike}
